---
title: "Julies_2022"
author: "Ry Dennis and Drew Anderson"
date: "5/18/2024"
output:
  word_document: default
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown of Analysis started 20220708 by Ry 
## Recent update 20240518 by Drew

## Libraries
```{r libraries}
library(lme4)
library(lmerTest)
library(dplyr)
library(ggplot2)
library(car)
```

## Data Entry and Curation

```{r data_entry}
#Data organizing
#####
#Prior to csv creation, I summed hitdivider.C and bite.C in excel to create aggression.C
df.clean<-read.csv("/Users/andersond/Desktop/Intruder_MS/Data and R Analysis/full_data.csv")
df<-df.clean

# round the states to remove decimal
df$closedivider.S <-  round(df$closedivider.S, digits = 0)
df$outnest.S <- round(df$outnest.S, digits = 0)
```

## Background organization for plotting developed by Ry
## Plot formatting changed at Journal Request for submission by Drew
## These plot commands adjust levels and therefore are relevent to the initial stats developed by Ry so this chunk remains unchanged

```{r plots}
df$order <- as.factor(df$order)
# code for stats and plots for Dennis et al 2022 manuscript
# use this line for published script and supply this data file.
# df <- read.csv(file = "Julies_2022_SuppleTable_BehaviorData.csv")  
df$plot.label <- as.factor(paste(df$pairtype,df$int.sex,df$sex, sep="."))
df$plot.label.order.D <- df$plot.label # makes a duplicate column
df$plot.label.order.D <- factor(df$plot.label, levels = levels(df$plot.label)[c(1,3,6,8,5,7,2,4)]) # reorders data and levels for labels

```

## Realization of a new stats to check during Ry analysis of intra-partner aggression. It's one-sided to large individuals so data table is cleaned up for this particular analysis

```{r}
#### <>< <>< <>< <>< <>< <>< Attack Mate with only large subjects  <><  ><> ><> ><> ><> ><> ><>  #### 
#### Make a new dataframe that only includes large fish ####
df.large <- df%>%
  filter(size=="large")

#I put 20230209_relevant_data.csv into Numbers for Mac and munged it to get a sum of mate attacks for each large fish for each intruder sex condition.
#trial number is no longer a factor because each attmate.sum is an aggregate of two trials.
df.large.AMSum<-read.csv("/Users/andersond/Desktop/Intruder_MS/Data and R Analysis/attack_mate.csv")

aov(attmate.sum ~sex*Int.sex, data=df.large.AMSum)
summary(aov(attmate.sum ~sex*Int.sex, data=df.large.AMSum))[[1]][1:3, 4:5]
```

## "Order" is a significant factor as seen in later analyses so Ry initially plotted a multipanel boxplot.
## Drew cleaned up the plot and converted response variables that use poisson distribution to log based plots

````{r}
#### order ####
par(mfrow=c(2,2),mar=c(0,6.5,.5,.5))
boxplot(df$aggression.C+1~df$order, xlab = "", xaxt="n", yaxt="n", ylab="Bites (#/10min)",log="y",cex.lab=2)
axis(2,at=c(1,6,21,101),label=c("0","5","20","100"),cex.axis=1.5)
par(mar=c(0,6.5,.5,.5))
boxplot(df$closedivider.S+1~df$order, xlab = "", xaxt="n",yaxt="n", ylab="Time (sec/10min)\n Close to Divider",log="y",cex.lab=2)
axis(2,at=c(1,11,101,501),label=c("0","10","100","500"),cex.axis=1.5)
par(mar=c(5,6.5,.5,.5))
boxplot(df$lateralroll.C+1~df$order, xlab = "Trial Number", yaxt="n",ylab="Lateral Rolls (#/10min)",log="y",cex.lab=2)
axis(2,at=c(1,3,6,21), label=c("0","2","5","20"),cex.axis=1.5)
par(mar=c(5,6.5,.5,.5))
boxplot(df$outnest.S~df$order, xlab = "Trial Number", ylab="Time (sec/10min)\nOut of Nest",cex.lab=2,cex.axis=1.5)

```


Set up GLMM tests with each condition as its own factor so that we can do pairwise comparisons.
For Poisson distributed values: aggression, close to divider, and lateral roll, optimize with bobyqa.
Gaussian distributed values were done with a lmer and not optimized.

## This analysis was perfomed by Ry and verified by supervising authors, Suzy Renn and Drew Anderson

```{r}
############################
## Aggression graphs linear estimate
############################
######
# vectors for plotting coded in df as $plot.label 
# add leading letters to make them plot in order. a, b, c, etc
#set levels this order for the first round of comparisons. 
#The intercept will be changed because in order to get the same test for all relevant comparisons, we need to change which value is the intercept. We cannot do linearHypothesis() with the intercept because comparing against the intercept gives meaningless results. this allows us to do the same test with all comparisons of interest.

levels(df$plot.label)
#first letter is intruder sex, second is subject sex, third is subject size
h.f.F.L <- df$plot.label == "femaleLarger.female.Female" #changed below
b.m.F.L <- df$plot.label == "femaleLarger.male.Female" 
#second term is intruder sex, third is subject sex
c.f.M.L <- df$plot.label == "maleLarger.female.Male"
d.m.M.L <- df$plot.label == "maleLarger.male.Male" 
e.f.F.S <- df$plot.label == "maleLarger.female.Female"
f.m.F.S <- df$plot.label == "maleLarger.male.Female"
g.f.M.S <- df$plot.label == "femaleLarger.female.Male"
a.m.M.S <- df$plot.label == "femaleLarger.male.Male" #making this the intercept because I am not comparing that to "femaleLarger.female.Female"

## create treatment codes
Codes <- matrix(0,nrow=8,ncol=8) #8 conditions
Codes[,1] <- rep(1,8)
Codes[2:8,2:8] <- diag(7)

#model 
glm2plot <- glmer(data = df, aggression.C~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.f.F.L + order + (1|ID)+(1|IntruderID),control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)),family="poisson") 
summary(glm2plot) 
#run pairwise comparisons to get 
#intruder sex comparison: large Female, female intruder vs large Female, male intruder  (b.m.F.L vs h.f.F.L)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 1, 0, 0, 0, 0, 0,-1, 0, 0, 0))  

#Subject sex comparison H.f.F.L vs C.F.M.L
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 1, 0 ,0, 0, 0,-1, 0, 0, 0))
#size comparison H.f.F.L vs e.f.F.S
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 0, 0 ,1, 0, 0,-1, 0, 0, 0))
#between mates H.f.F.L vs g.f.M.S
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 0, 0 ,0, 0, 1,-1, 0, 0, 0))
```

```{r}
#aggression
# change the order to do the rest of the comparisons
levels(df$plot.label)
a.f.F.L <- df$plot.label == "femaleLarger.female.Female"
b.m.F.L <- df$plot.label == "femaleLarger.male.Female"
c.f.M.L <- df$plot.label == "maleLarger.female.Male"
d.m.M.L <- df$plot.label == "maleLarger.male.Male"
e.f.F.S <- df$plot.label == "maleLarger.female.Female"
f.m.F.S <- df$plot.label == "maleLarger.male.Female"
g.f.M.S <- df$plot.label == "femaleLarger.female.Male"
h.m.M.S <- df$plot.label == "femaleLarger.male.Male"

## create treatment codes
Codes <- matrix(0,nrow=8,ncol=8) #8 conditions
Codes[,1] <- rep(1,8)
Codes[2:8,2:8] <- diag(7)

#model for pairwise comparisons
glm2plot <- glmer(data = df, aggression.C~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.m.M.S + order + (1|ID)+(1|IntruderID),control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)),family="poisson")

summary(glm2plot) 
# the following does the pairwise comparisons that aren't in the initial comparison

#intruder sex
# large Female, female intruder vs large Female, male intruder (a.f.F.L vs b.m.F.L) (see above)
# small male, female intruder vs small male male intruder (g vs h)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,0,0,1,-1,0,0,0))
# small female, female intruder vs small female male intruder (e vs f)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,1,-1,0,0,0,0,0))
#large male, female intruder vs large male male intruder (c vs d)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,-1,0,0,0,0,0,0,0))         

# Sex (larger subjects w female.int)  (a vs c) See above
# Sex (larger subjects w male.int)  (b vs d)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,-1,0,0,0,0,0,0,0))
# Sex (smaller w female.int) (e vs g)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,1,0,-1,0,0,0,0))
# Sex (smaller w male.int) (f vs h)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,0,1,0,-1,0,0,0))

# Size (Female w female.int) (a vs e) See above
# Size (Female w male.int) (b vs f)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,0,0,-1,0,0,0,0,0))
# Size (Male w female.int) (c vs g)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,0,0,0,-1,0,0,0,0))
# Size (Male w male.int) (d vs h)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,1,0,0,0,-1,0,0,0))

# Size and sex (femalelarger w female.int) (a vs g) see above
# Size and sex (Femalelarger w male.int) (b vs h)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,0,0,0,0,-1,0,0,0))
# Size and sex (Malelarger w female.int) (c vs e)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,0,-1,0,0,0,0,0,0)) 
# Size and sex (Malelarger w male.int) (d vs f)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,1,0,-1,0,0,0,0,0))


M <- vcov(glm2plot)[c(1:8),c(1:8)]
FE <- fixef(glm2plot)[c(1:8)]
Estimates <- as.numeric(Codes %*% FE)
EstVar <- rep(0,length(Estimates))
#
for(i in 1:8) EstVar[i] <- as.numeric(Codes[i,] %*% M %*% t(Codes)[,i])
EstSE <- sqrt(EstVar)
CI <- cbind(Estimates - 1.96*EstSE, Estimates + 1.96*EstSE) # is this fair if its not normal?


###Bit of code from Drew to save relevant data

Estimatedf<-data.frame(exp(Estimates))
Estimatedf$sex<-c("Female","Female","Male","Male","Female","Female","Male","Male")
Estimatedf$size<-c("large","large","large","large","small","small","small","small")
Estimatedf$int.sex<-c("Female","Male","Female","Male","Female","Male","Female","Male")
Estimatedf$sexsize<-paste0(Estimatedf$sex,Estimatedf$size)
Estimatedf$sexsize<-replace(Estimatedf$sexsize,Estimatedf$sexsize=="Femalelarge","Large Female Estimate")
Estimatedf$sexsize<-replace(Estimatedf$sexsize,Estimatedf$sexsize=="Malelarge","Large Male Estimate")
Estimatedf$sexsize<-replace(Estimatedf$sexsize,Estimatedf$sexsize=="Malesmall","Small Male Estimate")
Estimatedf$sexsize<-replace(Estimatedf$sexsize,Estimatedf$sexsize=="Femalesmall","Small Female Estimate")
names(Estimatedf)[names(Estimatedf) == 'sexsize'] <- 'Category'
names(Estimatedf)[names(Estimatedf) == 'exp.Estimates.'] <- 'Aggression'

```

```{r}
#close divider
######
levels(df$plot.label)
a.m.M.S <- df$plot.label == "femaleLarger.male.Male" #switch these below
b.m.F.L <- df$plot.label == "femaleLarger.male.Female"
c.f.M.L <- df$plot.label == "maleLarger.female.Male"
d.m.M.L <- df$plot.label == "maleLarger.male.Male" 
e.f.F.S <- df$plot.label == "maleLarger.female.Female"
f.m.F.S <- df$plot.label == "maleLarger.male.Female"
g.f.M.S <- df$plot.label == "femaleLarger.female.Male"
h.f.F.L <- df$plot.label == "femaleLarger.female.Female" #switch these below

## create treatment codes
Codes <- matrix(0,nrow=8,ncol=8) # 8 conditions
Codes[,1] <- rep(1,8)
Codes[2:8,2:8] <- diag(7)

####
glm2plot <- glmer(data = df, closedivider.S~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.f.F.L + order + (1|ID)+(1|IntruderID),control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)),family="poisson") 
summary(glm2plot) 
#h.f.F.L vs b.m.F.L    intruder sex                           
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 1, 0, 0, 0, 0, 0,-1, 0, 0, 0))  
#H.f.F.L vs C.F.M.L    subject sex     
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 1, 0 ,0, 0, 0,-1, 0, 0, 0))
#H.f.F.L vs e.f.F.S    size            
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 0, 0 ,1, 0, 0,-1, 0, 0, 0))
#H.f.F.L vs g.f.M.S    between mates   
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 0, 0 ,0, 0, 1,-1, 0, 0, 0))# change the order to do the rest of the comparisons
````

```{r}
############################
##### closedivider graphs linear estimate ##### 
levels(df$plot.label)
a.f.F.L <- df$plot.label == "femaleLarger.female.Female"
b.m.F.L <- df$plot.label == "femaleLarger.male.Female"
c.f.M.L <- df$plot.label == "maleLarger.female.Male"
d.m.M.L <- df$plot.label == "maleLarger.male.Male"
e.f.F.S <- df$plot.label == "maleLarger.female.Female"
f.m.F.S <- df$plot.label == "maleLarger.male.Female"
g.f.M.S <- df$plot.label == "femaleLarger.female.Male"
h.m.M.S <- df$plot.label == "femaleLarger.male.Male"

## create treatment codes
Codes <- matrix(0,nrow=8,ncol=8) #8 conditions
Codes[,1] <- rep(1,8)
Codes[2:8,2:8] <- diag(7)

#####
glm2plot <- glmer(data = df, df$closedivider.S~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.m.M.S + order+ (1|ID)+(1|IntruderID), control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)), family = "poisson") ## 

summary(glm2plot) 
# the following does the pairwise comparisons that aren't in the initial model
#two cyan (c vs d)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,-1,0,0,0,0,0,0,0))         
# two pink (e vs f) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,1,-1,0,0,0,0,0))
# two blue (g vs h)  
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,0,0,1,-1,0,0,0))
# two red (a vs b) 
#this value is from summary(glm2plot)

# Sex (larger w female.int)  (a vs c) see above
#this value is from summary(glm2plot)
# Sex (larger w male.int)  (b vs d) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,-1,0,0,0,0,0,0,0))
# Sex (smaller w female.int) (e vs g) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,1,0,-1,0,0,0,0))
# Sex (smaller w male.int) (f vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,0,1,0,-1,0,0,0))

# Size (Female w male.int) (b vs f) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,0,0,-1,0,0,0,0,0))
# Size (Male w female.int) (c vs g) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,0,0,0,-1,0,0,0,0))
# Size (Male w male.int) (d vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,1,0,0,0,-1,0,0,0))
# Size (Female w female.int) (a vs e) see above

##new
# Size and sex (femalelarger w female.int) (a vs g) is above
# Size and sex (Femalelarger w male.int) (b vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,0,0,0,0,-1,0,0,0))
# Size and sex (Malelarger w female.int) (c vs e) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,0,-1,0,0,0,0,0,0)) 
# Size and sex (Malelarger w male.int) (d vs f) (0.2875)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,1,0,-1,0,0,0,0,0))
```

```{r}
M <- vcov(glm2plot)[c(1:8),c(1:8)]
FE <- fixef(glm2plot)[c(1:8)]
Estimates <- as.numeric(Codes %*% FE)
EstVar <- rep(0,length(Estimates))
#
for(i in 1:8) EstVar[i] <- as.numeric(Codes[i,] %*% M %*% t(Codes)[,i])
EstSE <- sqrt(EstVar)


###Drew add-on
Estimatedf$Close<-exp(Estimates)
```

```{r}
#out of nest
# change the order to do the rest of the comparisons
######
levels(df$plot.label)
h.f.F.L <- df$plot.label == "femaleLarger.female.Female" #changed below
b.m.F.L <- df$plot.label == "femaleLarger.male.Female"
c.f.M.L <- df$plot.label == "maleLarger.female.Male"
d.m.M.L <- df$plot.label == "maleLarger.male.Male" 
e.f.F.S <- df$plot.label == "maleLarger.female.Female"
f.m.F.S <- df$plot.label == "maleLarger.male.Female"
g.f.M.S <- df$plot.label == "femaleLarger.female.Male"
a.m.M.S <- df$plot.label == "femaleLarger.male.Male"    #changed below

## create treatment codes
Codes <- matrix(0,nrow=8,ncol=8) #changed from 4 (old julies) to 8 cuz I have 8 conditions
Codes[,1] <- rep(1,8)
Codes[2:8,2:8] <- diag(7)

glm2plot <- glmer(data = df, outnest.S~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.f.F.L + order + (1|ID)+(1|IntruderID))
summary(glm2plot) 

#h.f.F.L vs b.m.F.L            #intruder sex              
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 1, 0, 0, 0, 0, 0,-1, 0, 0, 0))  
#H.f.F.L vs C.F.M.L     subject sex
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 1, 0 ,0, 0, 0,-1, 0, 0, 0))
#H.f.F.L vs e.f.F.S      subject size
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 0, 0 ,1, 0, 0,-1, 0, 0, 0))
#H.f.F.L vs g.f.M.S       between mates
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 0, 0 ,0, 0, 1,-1, 0, 0, 0))# change the order to do the rest of the comparisons
```
```{r}
############################
## out nest graphs linear estimate
############################
levels(df$plot.label)
a.f.F.L <- df$plot.label == "femaleLarger.female.Female"
b.m.F.L <- df$plot.label == "femaleLarger.male.Female"
c.f.M.L <- df$plot.label == "maleLarger.female.Male"
d.m.M.L <- df$plot.label == "maleLarger.male.Male"
e.f.F.S <- df$plot.label == "maleLarger.female.Female"
f.m.F.S <- df$plot.label == "maleLarger.male.Female"
g.f.M.S <- df$plot.label == "femaleLarger.female.Male"
h.m.M.S <- df$plot.label == "femaleLarger.male.Male"

## create treatment codes
Codes <- matrix(0,nrow=8,ncol=8) 
Codes[,1] <- rep(1,8)
Codes[2:8,2:8] <- diag(7)
#####
glm2plot <- lmer(data = df, outnest.S~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.m.M.S + order+ (1|ID)+(1|IntruderID)) 

summary(glm2plot) #(NS except order)
# the following does the pairwise comparisons that aren't in the initial model
# two red (a vs b) 
#two cyan (c vs d) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,-1,0,0,0,0,0,0,0))         
# two pink (e vs f) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,1,-1,0,0,0,0,0))
# two blue (g vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,0,0,1,-1,0,0,0))

# Sex (larger w female.int)  (a vs c) 
# Sex (larger w male.int)  (b vs d) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,-1,0,0,0,0,0,0,0))
# Sex (smaller w female.int) (e vs g) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,1,0,-1,0,0,0,0))
# Sex (smaller w male.int) (f vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,0,1,0,-1,0,0,0))

# Size (Female w female.int) (a vs e) see above
# Size (Female w male.int) (b vs f) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,0,0,-1,0,0,0,0,0))
# Size (Male w female.int) (c vs g) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,0,0,0,-1,0,0,0,0))
# Size (Male w male.int) (d vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,1,0,0,0,-1,0,0,0))

# Size and sex (femalelarger w female.int) (a vs g) see above
# Size and sex (Femalelarger w male.int) (b vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,0,0,0,0,-1,0,0,0))
# Size and sex (Malelarger w female.int) (c vs e) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,0,-1,0,0,0,0,0,0)) 
# Size and sex (Malelarger w male.int) (d vs f) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,1,0,-1,0,0,0,0,0))
```

```{r}
M <- vcov(glm2plot)[c(1:8),c(1:8)]
FE <- fixef(glm2plot)[c(1:8)]
Estimates <- as.numeric(Codes %*% FE)
EstVar <- rep(0,length(Estimates))
#
for(i in 1:8) EstVar[i] <- as.numeric(Codes[i,] %*% M %*% t(Codes)[,i])
EstSE <- sqrt(EstVar)
CI <- cbind(Estimates - 1.96*EstSE, Estimates + 1.96*EstSE) 

###Drew add-on
Estimatedf$Out<-Estimates
````

```{r}
# lateral roll
#change the order to do the rest of the comparisons
######
levels(df$plot.label)
h.f.F.L <- df$plot.label == "femaleLarger.female.Female" #changed below
b.m.F.L <- df$plot.label == "femaleLarger.male.Female"
c.f.M.L <- df$plot.label == "maleLarger.female.Male"
d.m.M.L <- df$plot.label == "maleLarger.male.Male" 
e.f.F.S <- df$plot.label == "maleLarger.female.Female"
f.m.F.S <- df$plot.label == "maleLarger.male.Female"
g.f.M.S <- df$plot.label == "femaleLarger.female.Male"
a.m.M.S <- df$plot.label == "femaleLarger.male.Male" #changed below

## create treatment codes
Codes <- matrix(0,nrow=8,ncol=8) 
Codes[,1] <- rep(1,8)
Codes[2:8,2:8] <- diag(7)


glm2plot <- glmer(data = df, lateralroll.C~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.f.F.L + order + (1|ID)+(1|IntruderID),control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)),family="poisson") 
summary(glm2plot) 

#h.f.F.L vs b.m.F.L          intruder sex                                       
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 1, 0, 0, 0, 0, 0,-1, 0, 0, 0))  
#H.f.F.L vs C.F.M.L           subject sex
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 1, 0 ,0, 0, 0,-1, 0, 0, 0))
#H.f.F.L vs e.f.F.S         subject size
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 0, 0 ,1, 0, 0,-1, 0, 0, 0))
#H.f.F.L vs g.f.M.S         between mates
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 0, 0 ,0, 0, 1,-1, 0, 0, 0))# change the order to do the rest of the comparisons
````

```{r}
## Lateral Roll graphs linear estimate
############################
levels(df$plot.label)
a.f.F.L <- df$plot.label == "femaleLarger.female.Female"
b.m.F.L <- df$plot.label == "femaleLarger.male.Female"
c.f.M.L <- df$plot.label == "maleLarger.female.Male"
d.m.M.L <- df$plot.label == "maleLarger.male.Male"
e.f.F.S <- df$plot.label == "maleLarger.female.Female"
f.m.F.S <- df$plot.label == "maleLarger.male.Female"
g.f.M.S <- df$plot.label == "femaleLarger.female.Male"
h.m.M.S <- df$plot.label == "femaleLarger.male.Male"

## create treatment codes
Codes <- matrix(0,nrow=8,ncol=8)
Codes[,1] <- rep(1,8)
Codes[2:8,2:8] <- diag(7)
# Codes
#####
glm2plot <- glmer(data = df, lateralroll.C~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.m.M.S + order+ (1|ID)+(1|IntruderID),control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)),family="poisson") 

summary(glm2plot) 
# the following does the pairwise comparisons that aren't in the initial model
# two red (a vs b) see above

# two blue (g vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,0,0,1,-1,0,0,0))
# two pink (e vs f) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,1,-1,0,0,0,0,0))
#two cyan (c vs d) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,-1,0,0,0,0,0,0,0))         

# Sex (larger w female.int)  (a vs c) 

# Sex (larger w male.int)  (b vs d) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,-1,0,0,0,0,0,0,0))
# Sex (smaller w female.int) (e vs g) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,1,0,-1,0,0,0,0))
# Sex (smaller w male.int) (f vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,0,1,0,-1,0,0,0))

# Size (Female w female.int) (a vs e) 

# Size (Female w male.int) (b vs f) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,0,0,-1,0,0,0,0,0))
# Size (Male w female.int) (c vs g) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,0,0,0,-1,0,0,0,0))
# Size (Male w male.int) (d vs h)  
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,1,0,0,0,-1,0,0,0))

##new
# Size and sex (femalelarger w female.int) (a vs g)  

# Size and sex (Femalelarger w male.int) (b vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,0,0,0,0,-1,0,0,0))
# Size and sex (Malelarger w female.int) (c vs e) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,0,-1,0,0,0,0,0,0)) 
# Size and sex (Malelarger w male.int) (d vs f) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,1,0,-1,0,0,0,0,0))
```

```{r}
M <- vcov(glm2plot)[c(1:8),c(1:8)]
FE <- fixef(glm2plot)[c(1:8)]
Estimates <- as.numeric(Codes %*% FE)
EstVar <- rep(0,length(Estimates))
#
for(i in 1:8) EstVar[i] <- as.numeric(Codes[i,] %*% M %*% t(Codes)[,i])
EstSE <- sqrt(EstVar)
#barplot(exp(Estimates), space = BarSpace, col=col.D, ylim=c(0, round(max(exp(Estimates)) + 0.1*max(exp(Estimates)),0)), ylab = "lateral roll", names.arg=names(FE), las = 2, cex.names = 0.5)

###Drew add-on
Estimatedf$Roll<-exp(Estimates)
```


##Drew Add on for graph

````{r}
df.clean$sexsize<-paste0(df.clean$sex,df.clean$size)
df.clean$int.sex<-replace(df.clean$int.sex,df.clean$int.sex=="female","Female")
df.clean$int.sex<-replace(df.clean$int.sex,df.clean$int.sex=="male","Male")
df.clean$sexsize<-replace(df.clean$sexsize,df.clean$sexsize=="Femalelarge","Large Female Raw")
df.clean$sexsize<-replace(df.clean$sexsize,df.clean$sexsize=="Malelarge","Large Male Raw")
df.clean$sexsize<-replace(df.clean$sexsize,df.clean$sexsize=="Malesmall","Small Male Raw")
df.clean$sexsize<-replace(df.clean$sexsize,df.clean$sexsize=="Femalesmall","Small Female Raw")
names(df.clean)[names(df.clean) == 'sexsize'] <- 'Category'
````

##Big plot
````{r}
library(gridExtra)
agg.plot<-ggplot(data=Estimatedf,aes(int.sex,Aggression))+geom_point(mapping=aes(colour=Category,shape=Category,size=Category))+geom_jitter(data=df.clean,aes(int.sex,(aggression.C+3)/8,colour = Category,shape=Category,size=Category),width = 0.19)+scale_y_log10(limits=c(0.3,26),"Estimated Bites (#/10min)",sec.axis=sec_axis(trans=~.*8,name="Raw Bites (#/10min)",breaks=c(3,13,33,103),labels=c("0","10","30","100")))+scale_shape_manual(values=c(16,1,17,2,16,1,17,2))+scale_size_manual(values=c(7,4,7,4,5,3,5,3))+scale_color_manual(values=c("#F26419","#F26419","#33658A","#33658A","#F6AE2D","#F6AE2D","#86BBD8","#86BBD8"))+geom_line(data=Estimatedf,aes(x=int.sex,y=Aggression,group = Category,linetype=Category,color = Category))+scale_linetype_manual(values=c(1,1,2,2))+theme(axis.text=element_text(size=12),axis.title=element_text(size=14,face="bold"),panel.background = element_rect(fill = 'white', colour = '#2F4858'))+scale_x_discrete("",labels=NULL)+guides(linetype="none")+theme(legend.position = "none",plot.margin = unit(c(.1,.5,0,.5),"inches"))

close.plot<-ggplot(data=Estimatedf,aes(int.sex,Close))+geom_point(mapping=aes(colour=Category,shape=Category,size=Category))+geom_jitter(data=df.clean,aes(int.sex,(closedivider.S+30)/3,colour = Category,shape=Category,size=Category),width = 0.19)+scale_y_log10(limits=c(9,210),"Estimated Time (sec/10min)\nClose to Divider",sec.axis=sec_axis(trans=~.*3,name="Raw Time (sec/10min)\nClose to Divider", breaks=c(30,130,330),labels=c("0","100","300")))+scale_shape_manual(values=c(16,1,17,2,16,1,17,2))+scale_size_manual(values=c(7,4,7,4,5,3,5,3))+scale_color_manual(values=c("#F26419","#F26419","#33658A","#33658A","#F6AE2D","#F6AE2D","#86BBD8","#86BBD8"))+geom_line(data=Estimatedf,aes(x=int.sex,y=Close,group = Category,linetype=Category,color = Category))+scale_linetype_manual(values=c(1,1,2,2))+theme(axis.text=element_text(size=12),axis.title=element_text(size=14,face="bold"),panel.background = element_rect(fill = 'white', colour = '#2F4858'))+scale_x_discrete("",labels=NULL)+guides(linetype="none")+theme(legend.position = "none",plot.margin = unit(c(.1,.5,0,.5),"inches"))

out.plot<-ggplot(data=Estimatedf,aes(int.sex,Out))+geom_point(mapping=aes(colour=Category,shape=Category,size=Category))+geom_jitter(data=df.clean,aes(int.sex,(outnest.S)/2,colour = Category,shape=Category,size=Category),width = 0.19)+scale_y_continuous(limits=c(-1,350),"Estimated Time (sec/10min)\nOut of Nest",sec.axis=sec_axis(trans=~.*2,name="Raw Time (sec/10min)\nOut of Nest"))+scale_shape_manual(values=c(16,1,17,2,16,1,17,2))+scale_size_manual(values=c(7,4,7,4,5,3,5,3))+scale_color_manual(values=c("#F26419","#F26419","#33658A","#33658A","#F6AE2D","#F6AE2D","#86BBD8","#86BBD8"))+geom_line(data=Estimatedf,aes(x=int.sex,y=Out,group = Category,linetype=Category,color = Category))+scale_linetype_manual(values=c(1,1,2,2))+theme(axis.text=element_text(size=12),axis.title=element_text(size=14,face="bold"),panel.background = element_rect(fill = 'white', colour = '#2F4858'))+scale_x_discrete("Intruder Sex")+guides(linetype="none")+theme(legend.position = "bottom",plot.margin = unit(c(.1,.5,0,.5),"inches"))

roll.plot<- ggplot(data=Estimatedf,aes(int.sex,Roll))+geom_point(mapping=aes(colour=Category,shape=Category,size=Category))+geom_jitter(data=df.clean,aes(int.sex,(lateralroll.C+1)/10,colour = Category,shape=Category,size=Category),width = 0.19)+scale_y_log10(limits=c(.05,3),"Estimated Lateral Rolls (#/10min)",sec.axis=sec_axis(trans=~.*10,name="Raw Lateral Rolls (#/10min)",breaks=c(1,4,11,31),labels=c("0","3","10","30")))+scale_shape_manual(values=c(16,1,17,2,16,1,17,2))+scale_size_manual(values=c(7,4,7,4,5,3,5,3))+scale_color_manual(values=c("#F26419","#F26419","#33658A","#33658A","#F6AE2D","#F6AE2D","#86BBD8","#86BBD8"))+geom_line(data=Estimatedf,aes(x=int.sex,y=Roll,group = Category,linetype=Category,color = Category))+scale_linetype_manual(values=c(1,1,2,2))+theme(axis.text=element_text(size=12),axis.title=element_text(size=14,face="bold"),panel.background = element_rect(fill = 'white', colour = '#2F4858'))+scale_x_discrete("Intruder Sex")+guides(linetype="none")+theme(legend.position = "none",plot.margin = unit(c(.1,.5,0,.5),"inches"))

grid.arrange(agg.plot,close.plot,roll.plot, out.plot,nrow = 2)
````

## Drew Add-on to investigate the aggression of the intruder as requested by initial reviewers. This code is the finalized version to test post-hoc pairwise comparisons

````{r}
newdf<-df[which(df$size=="large"),]
newdf$int.label<-as.factor(paste0(newdf$pairtype,".",newdf$int.sex))
mod7<-glmer(Int.attack~int.label+aggression.C+(1|IntruderID)+(1|pairID),family="poisson",newdf)
newdf$int.label<-factor(newdf$int.label, levels=c("femaleLarger.male","maleLarger.female","maleLarger.male","femaleLarger.female"))
mod8<-glmer(Int.attack~int.label+aggression.C+(1|IntruderID)+(1|pairID),family="poisson",newdf)
newdf$int.label<-factor(newdf$int.label, levels=c("maleLarger.female","maleLarger.male","femaleLarger.female","femaleLarger.male"))
mod9<-glmer(Int.attack~int.label+aggression.C+(1|IntruderID)+(1|pairID),family="poisson",newdf)
summary(mod7)
summary(mod8)
summary(mod9)
````

            ML.male   ML.female   FL.male   FL.female

ML.male     --

ML.female   0.0123    --

FL.male     0.0035    0.8780      --

FL.female   0.5285    0.0457      0.0199    --

So... intruders reciprocate aggression from the larger fish. And are more aggressive to opposite sex intruders!
## Plot to show this is true

````{r}
plot(Int.attack~int.label+aggression.C,newdf)
````

## Using Ry's code to set up terms to plot for my own plotting technique

```{r}
Codes <- matrix(0,nrow=4,ncol=4) #8 conditions
Codes[,1] <- rep(1,4)
Codes[2:4,2:4] <- diag(3)
M <- vcov(mod9)[c(1:4),c(1:4)]
FE <- fixef(mod9)[c(1:4)]
Estimates <- as.numeric(Codes %*% FE)
EstVar <- rep(0,length(Estimates))
#
for(i in 1:4) EstVar[i] <- as.numeric(Codes[i,] %*% M %*% t(Codes)[,i])
EstSE <- sqrt(EstVar)
 
###Bit of code from Drew to save relevant data
Intdf<-data.frame(exp(Estimates))
Intdf$PairType<-c("Male Larger","Male Larger","Female Larger","Female Larger")
Intdf$Intruder<-c("Estimated Female","Estimated Male","Estimated Female","Estimated Male")
names(Intdf)[names(Intdf) == 'exp.Estimates.'] <- 'IntruderBites'
newdf$pairtype<-replace(newdf$pairtype,newdf$pairtype=="femaleLarger","Female Larger")
newdf$pairtype<-replace(newdf$pairtype,newdf$pairtype=="maleLarger","Male Larger")
newdf$int.sex<-replace(newdf$int.sex,newdf$int.sex=="female","Raw Female")
newdf$int.sex<-replace(newdf$int.sex,newdf$int.sex=="male","Raw Male")
names(newdf)[names(newdf) == 'int.sex'] <- 'Intruder'

````

## Here's the plot

````{r}
ggplot(data=Intdf,aes(PairType,IntruderBites))+geom_point(mapping=aes(colour=Intruder,shape=Intruder,size=Intruder))+geom_jitter(data=newdf,aes(pairtype,(Int.attack+15)/5,colour = Intruder,shape=Intruder,size=Intruder),width = 0.19)+scale_y_log10(limits=c(3,16),"Estimated Intruder Bites (#/10min)",sec.axis=sec_axis(trans=~.*5,name="Raw Intruder Bites (#/10min)",breaks=c(15,25,40,75),labels=c("0","10","25","60")))+scale_shape_manual(values=c(16,17,1,2))+scale_size_manual(values=c(7,7,4,4))+scale_color_manual(values=c("#C89933","#694966","#C89933","#694966"))+geom_line(data=Intdf,aes(x=PairType,y=IntruderBites,group = Intruder,linetype=Intruder,color = Intruder))+scale_linetype_manual(values=c(1,2))+theme(axis.text=element_text(size=12),axis.title=element_text(size=14,face="bold"),panel.background = element_rect(fill = 'white', colour = '#2F4858'))+scale_x_discrete("Pair Type")+guides(linetype="none")+theme(legend.position = "bottom",plot.margin = unit(c(.1,.5,0,.5),"inches"))

````
