---
title: "Julie Intruder Study"
author: "Andrew Anderson and Ry Dennis Combined"
date: "2023-08-06"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r libraries}
library(lme4)
library(lmerTest)
library(dplyr)
library(ggplot2)
library(car)
```

## Data Entry

```{r data_entry}
#Data organizing
#####
#Prior to csv creation, I summed hitdivider.C and bite.C in excel to create aggression.C
df.clean<-read.csv("~/full_data.csv")

# round the states to remove decimal
df$closedivider.S <-  round(df$closedivider.S, digits = 0)
df$outnest.S <- round(df$outnest.S, digits = 0)
```

## Run the Plot
### These are the analyses done by Ry Dennis
```{r plots}
df$order <- as.factor(df$order)
# code for stats and plots for Dennis et al 2022 manuscript
# use this line for published script and supply this data file.
# df <- read.csv(file = "Julies_2022_SuppleTable_BehaviorData.csv")  
df$plot.label <- as.factor(paste(df$pairtype,df$int.sex,df$sex, sep="."))
df$plot.label.order.D <- df$plot.label # makes a duplicate column
df$plot.label.order.D <- factor(df$plot.label, levels = levels(df$plot.label)[c(1,3,6,8,5,7,2,4)]) # reorders data and levels for labels

spacing1 <- c(0,1,3,4,7,8,10,11) # used to place boxes in boxplot and data points
spacing2 <- c(0.5,3.5,7.5,10.5) # used for labeling x-axis

col.D = c("red","red","cyan","cyan","pink","pink", "blue","blue") # colors for plot

par(mar = c(10, 5.1, 4.1, 2.1)) 
## The default value for mar is c(5.1, 4.1, 4.1, 2.1) to reset it
```

```{r}
#### <>< <>< <>< <>< <>< <>< Attack Mate with only large subjects  <><  ><> ><> ><> ><> ><> ><>  #### 
#### Make a new dataframe that only includes large fish ####
df.large <- df%>%
  filter(size=="large")

#I put 20230209_relevant_data.csv into Numbers for Mac and munged it to get a sum of mate attacks for each large fish for each intruder sex condition.
#trial number is no longer a factor because each attmate.sum is an aggregate of two trials.
df.large.AMSum<-read.csv("~/attack_mate.csv")

aov(attmate.sum ~sex*Int.sex, data=df.large.AMSum)
summary(aov(attmate.sum ~sex*Int.sex, data=df.large.AMSum))[[1]][1:3, 4:5]
```

```{r}
##### Do stats with three-way model, order as fixed and ID and intruder ID as random ####
mod.aggression.C <- glmer(aggression.C~size*sex*int.sex+order+(1|ID)+(1|IntruderID), family = "poisson", data = df)
summary(mod.aggression.C) 

##### within body length of divider ####
# ( glmer it was significant sizeXsex ; sexXint.sex ; sizeXsexXint.sex but this is a continuous variable) (order as fixed or random doesn't change much)
mod.closedivider.S <- glmer(closedivider.S~size*sex*int.sex+order+(1|ID)+(1|IntruderID), data = df, family = "poisson") ### 
summary(mod.closedivider.S) 

# out of nest mutually exclusive with in nest
#### NS (with order as either fixed or random)####
mod.outnest.S <-       lmer(outnest.S~size*sex*int.sex+order+(1|ID)+(1|IntruderID), data = df)
summary(mod.outnest.S) 

##### Lateral roll####
mod.lateralroll.C <- glmer(lateralroll.C~size*sex*int.sex+order+(1|ID)+(1|IntruderID), control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)), family="poisson", data = df)  ## 
summary (mod.lateralroll.C)  

```

```{r}
############################
## Aggression graphs linear estimate
############################
######
# vectors for plotting coded in df as $plot.label 
# add leading letters to make them plot in order. a, b, c, etc
#set levels this order for the first round of comparisons. 
#The intercept will be changed because in order to get the same test for all relevant comparisons, we need to change which value is the intercept. We cannot do linearHypothesis() with the intercept because comparing against the intercept gives meaningless results. this allows us to do the same test with all comparisons of interest.

levels(df$plot.label)
#first letter is intruder sex, second is subject sex, third is subject size
h.f.F.L <- df$plot.label == "femaleLarger.female.Female" 
b.m.F.L <- df$plot.label == "femaleLarger.male.Female" 

#second term is intruder sex, third is subject sex

c.f.M.L <- df$plot.label == "maleLarger.female.Male"
d.m.M.L <- df$plot.label == "maleLarger.male.Male" 
e.f.F.S <- df$plot.label == "maleLarger.female.Female"
f.m.F.S <- df$plot.label == "maleLarger.male.Female"
g.f.M.S <- df$plot.label == "femaleLarger.female.Male"
a.m.M.S <- df$plot.label == "femaleLarger.male.Male" 

#making this the intercept because I am not comparing that to "femaleLarger.female.Female"

#model 
glm2plot <- glmer(data = df, aggression.C~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.f.F.L + order + (1|ID)+(1|IntruderID),control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)),family="poisson") 
summary(glm2plot) 
#run pairwise comparisons to get 
#intruder sex comparison: large Female, female intruder vs large Female, male intruder  (b.m.F.L vs h.f.F.L)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 1, 0, 0, 0, 0, 0,-1, 0, 0, 0))  

#Subject sex comparison H.f.F.L vs C.F.M.L
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 1, 0 ,0, 0, 0,-1, 0, 0, 0))
#size comparison H.f.F.L vs e.f.F.S
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 0, 0 ,1, 0, 0,-1, 0, 0, 0))
#between mates H.f.F.L vs g.f.M.S
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 0, 0 ,0, 0, 1,-1, 0, 0, 0))

#aggression
# change the order to do the rest of the comparisons
levels(df$plot.label)
a.f.F.L <- df$plot.label == "femaleLarger.female.Female"
b.m.F.L <- df$plot.label == "femaleLarger.male.Female"
c.f.M.L <- df$plot.label == "maleLarger.female.Male"
d.m.M.L <- df$plot.label == "maleLarger.male.Male"
e.f.F.S <- df$plot.label == "maleLarger.female.Female"
f.m.F.S <- df$plot.label == "maleLarger.male.Female"
g.f.M.S <- df$plot.label == "femaleLarger.female.Male"
h.m.M.S <- df$plot.label == "femaleLarger.male.Male"

#model for pairwise comparisons
glm2plot <- glmer(data = df, aggression.C~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.m.M.S + order + (1|ID)+(1|IntruderID),control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)),family="poisson")

summary(glm2plot) 
# the following does the pairwise comparisons that aren't in the initial comparison

#intruder sex
# large Female, female intruder vs large Female, male intruder (a.f.F.L vs b.m.F.L) (see above)
# small male, female intruder vs small male male intruder (g vs h)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,0,0,1,-1,0,0,0))
# small female, female intruder vs small female male intruder (e vs f)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,1,-1,0,0,0,0,0))
#large male, female intruder vs large male male intruder (c vs d)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,-1,0,0,0,0,0,0,0))         

# Sex (larger subjects w female.int)  (a vs c) See above
# Sex (larger subjects w male.int)  (b vs d)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,-1,0,0,0,0,0,0,0))
# Sex (smaller w female.int) (e vs g)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,1,0,-1,0,0,0,0))
# Sex (smaller w male.int) (f vs h)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,0,1,0,-1,0,0,0))

# Size (Female w female.int) (a vs e) See above
# Size (Female w male.int) (b vs f)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,0,0,-1,0,0,0,0,0))
# Size (Male w female.int) (c vs g)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,0,0,0,-1,0,0,0,0))
# Size (Male w male.int) (d vs h)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,1,0,0,0,-1,0,0,0))

# Size and sex (femalelarger w female.int) (a vs g) see above
# Size and sex (Femalelarger w male.int) (b vs h)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,0,0,0,0,-1,0,0,0))
# Size and sex (Malelarger w female.int) (c vs e)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,0,-1,0,0,0,0,0,0)) 
# Size and sex (Malelarger w male.int) (d vs f)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,1,0,-1,0,0,0,0,0))
````

```{r}
#close divider

levels(df$plot.label)
a.m.M.S <- df$plot.label == "femaleLarger.male.Male" 
b.m.F.L <- df$plot.label == "femaleLarger.male.Female"
c.f.M.L <- df$plot.label == "maleLarger.female.Male"
d.m.M.L <- df$plot.label == "maleLarger.male.Male" 
e.f.F.S <- df$plot.label == "maleLarger.female.Female"
f.m.F.S <- df$plot.label == "maleLarger.male.Female"
g.f.M.S <- df$plot.label == "femaleLarger.female.Male"
h.f.F.L <- df$plot.label == "femaleLarger.female.Female" 

glm2plot <- glmer(data = df, closedivider.S~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.f.F.L + order + (1|ID)+(1|IntruderID),control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)),family="poisson") 
summary(glm2plot) 
#h.f.F.L vs b.m.F.L    intruder sex                           
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 1, 0, 0, 0, 0, 0,-1, 0, 0, 0))  
#H.f.F.L vs C.F.M.L    subject sex     
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 1, 0 ,0, 0, 0,-1, 0, 0, 0))
#H.f.F.L vs e.f.F.S    size            
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 0, 0 ,1, 0, 0,-1, 0, 0, 0))
#H.f.F.L vs g.f.M.S    between mates   
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 0, 0 ,0, 0, 1,-1, 0, 0, 0))

# change the order to do the rest of the comparisons

levels(df$plot.label)
a.f.F.L <- df$plot.label == "femaleLarger.female.Female"
b.m.F.L <- df$plot.label == "femaleLarger.male.Female"
c.f.M.L <- df$plot.label == "maleLarger.female.Male"
d.m.M.L <- df$plot.label == "maleLarger.male.Male"
e.f.F.S <- df$plot.label == "maleLarger.female.Female"
f.m.F.S <- df$plot.label == "maleLarger.male.Female"
g.f.M.S <- df$plot.label == "femaleLarger.female.Male"
h.m.M.S <- df$plot.label == "femaleLarger.male.Male"

glm2plot <- glmer(data = df, df$closedivider.S~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.m.M.S + order+ (1|ID)+(1|IntruderID), control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)), family = "poisson") ## 

summary(glm2plot) 
# the following does the pairwise comparisons that aren't in the initial model
#two cyan (c vs d)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,-1,0,0,0,0,0,0,0))         
# two pink (e vs f) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,1,-1,0,0,0,0,0))
# two blue (g vs h)  
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,0,0,1,-1,0,0,0))
# two red (a vs b) 
#this value is from summary(glm2plot)

# Sex (larger w female.int)  (a vs c) see above
#this value is from summary(glm2plot)
# Sex (larger w male.int)  (b vs d) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,-1,0,0,0,0,0,0,0))
# Sex (smaller w female.int) (e vs g) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,1,0,-1,0,0,0,0))
# Sex (smaller w male.int) (f vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,0,1,0,-1,0,0,0))

# Size (Female w male.int) (b vs f) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,0,0,-1,0,0,0,0,0))
# Size (Male w female.int) (c vs g) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,0,0,0,-1,0,0,0,0))
# Size (Male w male.int) (d vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,1,0,0,0,-1,0,0,0))
# Size (Female w female.int) (a vs e) see above

##new
# Size and sex (femalelarger w female.int) (a vs g) is above
# Size and sex (Femalelarger w male.int) (b vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,0,0,0,0,-1,0,0,0))
# Size and sex (Malelarger w female.int) (c vs e) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,0,-1,0,0,0,0,0,0)) 
# Size and sex (Malelarger w male.int) (d vs f) (0.2875)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,1,0,-1,0,0,0,0,0))
```

```{r}
#out of nest
# change the order to do comparisons
######
levels(df$plot.label)
h.f.F.L <- df$plot.label == "femaleLarger.female.Female" 
b.m.F.L <- df$plot.label == "femaleLarger.male.Female"
c.f.M.L <- df$plot.label == "maleLarger.female.Male"
d.m.M.L <- df$plot.label == "maleLarger.male.Male" 
e.f.F.S <- df$plot.label == "maleLarger.female.Female"
f.m.F.S <- df$plot.label == "maleLarger.male.Female"
g.f.M.S <- df$plot.label == "femaleLarger.female.Male"
a.m.M.S <- df$plot.label == "femaleLarger.male.Male"    #changed below

glm2plot <- glmer(data = df, outnest.S~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.f.F.L + order + (1|ID)+(1|IntruderID))
summary(glm2plot) 

#h.f.F.L vs b.m.F.L            #intruder sex              
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 1, 0, 0, 0, 0, 0,-1, 0, 0, 0))  
#H.f.F.L vs C.F.M.L     subject sex
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 1, 0 ,0, 0, 0,-1, 0, 0, 0))
#H.f.F.L vs e.f.F.S      subject size
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 0, 0 ,1, 0, 0,-1, 0, 0, 0))
#H.f.F.L vs g.f.M.S       between mates
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 0, 0 ,0, 0, 1,-1, 0, 0, 0))

# change the order to do the rest of the comparisons
levels(df$plot.label)
a.f.F.L <- df$plot.label == "femaleLarger.female.Female"
b.m.F.L <- df$plot.label == "femaleLarger.male.Female"
c.f.M.L <- df$plot.label == "maleLarger.female.Male"
d.m.M.L <- df$plot.label == "maleLarger.male.Male"
e.f.F.S <- df$plot.label == "maleLarger.female.Female"
f.m.F.S <- df$plot.label == "maleLarger.male.Female"
g.f.M.S <- df$plot.label == "femaleLarger.female.Male"
h.m.M.S <- df$plot.label == "femaleLarger.male.Male"

glm2plot <- lmer(data = df, outnest.S~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.m.M.S + order+ (1|ID)+(1|IntruderID)) 

summary(glm2plot) #(NS except order)
# the following does the pairwise comparisons that aren't in the initial model
# two red (a vs b) 
#two cyan (c vs d) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,-1,0,0,0,0,0,0,0))         
# two pink (e vs f) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,1,-1,0,0,0,0,0))
# two blue (g vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,0,0,1,-1,0,0,0))

# Sex (larger w female.int)  (a vs c) 
# Sex (larger w male.int)  (b vs d) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,-1,0,0,0,0,0,0,0))
# Sex (smaller w female.int) (e vs g) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,1,0,-1,0,0,0,0))
# Sex (smaller w male.int) (f vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,0,1,0,-1,0,0,0))

# Size (Female w female.int) (a vs e) see above
# Size (Female w male.int) (b vs f) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,0,0,-1,0,0,0,0,0))
# Size (Male w female.int) (c vs g) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,0,0,0,-1,0,0,0,0))
# Size (Male w male.int) (d vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,1,0,0,0,-1,0,0,0))

# Size and sex (femalelarger w female.int) (a vs g) see above
# Size and sex (Femalelarger w male.int) (b vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,0,0,0,0,-1,0,0,0))
# Size and sex (Malelarger w female.int) (c vs e) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,0,-1,0,0,0,0,0,0)) 
# Size and sex (Malelarger w male.int) (d vs f) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,1,0,-1,0,0,0,0,0))
```

```{r}
# lateral roll
#change the order to do comparisons

levels(df$plot.label)
h.f.F.L <- df$plot.label == "femaleLarger.female.Female" 
b.m.F.L <- df$plot.label == "femaleLarger.male.Female"
c.f.M.L <- df$plot.label == "maleLarger.female.Male"
d.m.M.L <- df$plot.label == "maleLarger.male.Male" 
e.f.F.S <- df$plot.label == "maleLarger.female.Female"
f.m.F.S <- df$plot.label == "maleLarger.male.Female"
g.f.M.S <- df$plot.label == "femaleLarger.female.Male"
a.m.M.S <- df$plot.label == "femaleLarger.male.Male" #changed below

glm2plot <- glmer(data = df, lateralroll.C~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.f.F.L + order + (1|ID)+(1|IntruderID),control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)),family="poisson") 
summary(glm2plot) 

#h.f.F.L vs b.m.F.L          intruder sex                           
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 1, 0, 0, 0, 0, 0,-1, 0, 0, 0))  
#H.f.F.L vs C.F.M.L           subject sex
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 1, 0 ,0, 0, 0,-1, 0, 0, 0))
#H.f.F.L vs e.f.F.S         subject size
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 0, 0 ,1, 0, 0,-1, 0, 0, 0))
#H.f.F.L vs g.f.M.S         between mates
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0, 0, 0, 0 ,0, 0, 1,-1, 0, 0, 0))

# change the order to do more comparisons
levels(df$plot.label)
a.f.F.L <- df$plot.label == "femaleLarger.female.Female"
b.m.F.L <- df$plot.label == "femaleLarger.male.Female"
c.f.M.L <- df$plot.label == "maleLarger.female.Male"
d.m.M.L <- df$plot.label == "maleLarger.male.Male"
e.f.F.S <- df$plot.label == "maleLarger.female.Female"
f.m.F.S <- df$plot.label == "maleLarger.male.Female"
g.f.M.S <- df$plot.label == "femaleLarger.female.Male"
h.m.M.S <- df$plot.label == "femaleLarger.male.Male"

glm2plot <- glmer(data = df, lateralroll.C~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.m.M.S + order+ (1|ID)+(1|IntruderID),control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)),family="poisson") 

summary(glm2plot) 
# the following does the pairwise comparisons that aren't in the initial model
# two red (a vs b) see above

# two blue (g vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,0,0,1,-1,0,0,0))
# two pink (e vs f) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,1,-1,0,0,0,0,0))
#two cyan (c vs d) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,-1,0,0,0,0,0,0,0))         

# Sex (larger w female.int)  (a vs c) 

# Sex (larger w male.int)  (b vs d) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,-1,0,0,0,0,0,0,0))
# Sex (smaller w female.int) (e vs g) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,1,0,-1,0,0,0,0))
# Sex (smaller w male.int) (f vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,0,1,0,-1,0,0,0))

# Size (Female w female.int) (a vs e) 

# Size (Female w male.int) (b vs f) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,0,0,-1,0,0,0,0,0))
# Size (Male w female.int) (c vs g) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,0,0,0,-1,0,0,0,0))
# Size (Male w male.int) (d vs h)  
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,1,0,0,0,-1,0,0,0))

##new
# Size and sex (femalelarger w female.int) (a vs g)  

# Size and sex (Femalelarger w male.int) (b vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,0,0,0,0,-1,0,0,0))
# Size and sex (Malelarger w female.int) (c vs e) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,0,-1,0,0,0,0,0,0)) 
# Size and sex (Malelarger w male.int) (d vs f) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,1,0,-1,0,0,0,0,0))
```


###This is where Drew took over for graphs and analysis
```{r}
#### order ####
setEPS()
postscript("~/order.eps",print.it = TRUE)
par(mfrow=c(2,2),oma=c(5,1,1,1),mar=c(0,5,1,1)) 
boxplot(df$aggression.C~df$order, cex.axis=1.25, cex.lab=1.5,xaxt="n", xlab="",ylab="Bites (#/10min)")
boxplot(df$closedivider.S~df$order, cex.axis=1.25, cex.lab=1.5, xaxt="n", xlab="",ylab="Close to divider (s/10min)")
boxplot(df$outnest.S~df$order, cex.axis=1.25, cex.lab =1.5,xlab = "", ylab="Out of Nest (s/10min)")
boxplot(df$lateralroll.C~df$order, cex.axis=1.25,cex.lab=1.5,xlab = "", ylab="Lateral Roll (#/10min)")
title(xlab="Trial Number",outer=TRUE,cex.lab=2)
dev.off()
#NOT USED
#boxplot(df.large$attmate.C~df.large$order, xlab = "Trial Number", ylab="Attack Mate")
```


```{r}
df$order <- as.factor(df$order)
# code for stats and plots for Dennis et al 2022 manuscript
# use this line for published script and supply this data file.
# df <- read.csv(file = "Julies_2022_SuppleTable_BehaviorData.csv")  
df$plot.label <- as.factor(paste(df$pairtype,df$int.sex,df$sex, sep="."))
df$plot.label.order.D <- df$plot.label # makes a duplicate column
df$plot.label.order.D <- factor(df$plot.label, levels = levels(df$plot.label)[c(1,3,6,8,5,7,2,4)]) # reorders data and levels for labels

# change the order to do the rest of the comparisons
levels(df$plot.label)
a.f.F.L <- df$plot.label == "femaleLarger.female.Female"
b.m.F.L <- df$plot.label == "femaleLarger.male.Female"
c.f.M.L <- df$plot.label == "maleLarger.female.Male"
d.m.M.L <- df$plot.label == "maleLarger.male.Male"
e.f.F.S <- df$plot.label == "maleLarger.female.Female"
f.m.F.S <- df$plot.label == "maleLarger.male.Female"
g.f.M.S <- df$plot.label == "femaleLarger.female.Male"
h.m.M.S <- df$plot.label == "femaleLarger.male.Male"

## create treatment codes
Codes <- matrix(0,nrow=8,ncol=8) #8 conditions
Codes[,1] <- rep(1,8)
Codes[2:8,2:8] <- diag(7)
# Codes
col.D = c("red","red","cyan","cyan","pink","pink", "blue","blue")
BarSpace <- c(0.5,1,2,1,3,1,2,1) #redo the codes
#model for pairwise comparisons

glm2plotagg <- glmer(data = df, aggression.C~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.m.M.S + order + (1|ID)+(1|IntruderID),control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)),family="poisson")

Magg <- vcov(glm2plotagg)[c(1:8),c(1:8)]
FEagg <- fixef(glm2plotagg)[c(1:8)]
Estimatesagg <- as.numeric(Codes %*% FEagg)
EstVaragg <- rep(0,length(Estimatesagg))
#
for(i in 1:8) EstVaragg[i] <- as.numeric(Codes[i,] %*% Magg %*% t(Codes)[,i])
EstSEagg <- sqrt(EstVaragg)

glm2plotclose <- glmer(data = df, df$closedivider.S~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.m.M.S + order+ (1|ID)+(1|IntruderID), control= glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=2e5)), family = "poisson") ## 
Mclose <- vcov(glm2plotclose)[c(1:8),c(1:8)]
FEclose <- fixef(glm2plotclose)[c(1:8)]
Estimatesclose <- as.numeric(Codes %*% FEclose)
EstVarclose <- rep(0,length(Estimatesclose))
#
for(i in 1:8) EstVarclose[i] <- as.numeric(Codes[i,] %*% Mclose %*% t(Codes)[,i])
EstSEclose <- sqrt(EstVarclose)

glm2plotout <- lmer(data = df, outnest.S~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.m.M.S + order+ (1|ID)+(1|IntruderID)) 

Mout <- vcov(glm2plotout)[c(1:8),c(1:8)]
FEout <- fixef(glm2plotout)[c(1:8)]
Estimatesout <- as.numeric(Codes %*% FEout)
EstVarout <- rep(0,length(Estimatesout))
#
for(i in 1:8) EstVarout[i] <- as.numeric(Codes[i,] %*% Mout %*% t(Codes)[,i])
EstSEout <- sqrt(EstVarout)

glm2plotroll <- glmer(data = df, df$lateralroll.C~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.m.M.S + order+ (1|ID)+(1|IntruderID), control= glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=2e5)), family = "poisson") ## 
Mroll <- vcov(glm2plotroll)[c(1:8),c(1:8)]
FEroll <- fixef(glm2plotroll)[c(1:8)]
Estimatesroll <- as.numeric(Codes %*% FEroll)
EstVarroll <- rep(0,length(Estimatesroll))
#
for(i in 1:8) EstVarroll[i] <- as.numeric(Codes[i,] %*% Mroll %*% t(Codes)[,i])
EstSEroll <- sqrt(EstVarroll)
````

````{r}
setEPS()
postscript("~/all4.eps",print.it = TRUE)

par(mfrow=c(2,2),oma=c(9,1,1,1),mar=c(0,5,.5,0)) 

plot(rep(c(1,2),4), exp(Estimatesagg),  pch = c(21,21,23,23,21,21,23,23), cex = c(2, 2, 2, 2, 1, 1, 1, 1), ylim=c(0, round(max(exp(Estimatesagg)) + 0.1*max(exp(Estimatesagg)),0)), cex.axis=1.25, cex.lab=1.5, xlim = c(0.75,2.25), xlab = "", xaxt = "n", ylab = "Bites (#/10min)", bg= col.D, col = "black")
# ? lines # there must be a better way but this hack will work
lines( x= c(1,2), y = exp(Estimatesagg[c(1,2)]), col = "red", lty = 1,lwd=2)
lines( x= c(1,2), y = exp(Estimatesagg[c(3,4)]), col = "cyan", lty = 2,lwd=2)
lines( x= c(1,2), y = exp(Estimatesagg[c(5,6)]), col = "pink", lty = 2,lwd=2)
lines( x= c(1,2), y = exp(Estimatesagg[c(7,8)]), col = "blue", lty = 1,lwd =2)

plot(rep(c(1,2),4), exp(Estimatesclose),  pch = c(21,21,23,23,21,21,23,23), cex = c(2, 2, 2, 2, 1, 1, 1, 1), ylim=c(0, round(max(exp(Estimatesclose)) + 0.1*max(exp(Estimatesclose)),0)), cex.axis=1.25, cex.lab=1.25, xlim = c(0.75,2.25), xlab = "", xaxt = "n", ylab = "Close to Divider (s/10min)", bg= col.D, col = "black")
# ? lines # there must be a better way but this hack will work
lines( x= c(1,2), y = exp(Estimatesclose[c(1,2)]), col = "red", lty = 1,lwd=2)
lines( x= c(1,2), y = exp(Estimatesclose[c(3,4)]), col = "cyan", lty = 2,lwd=2)
lines( x= c(1,2), y = exp(Estimatesclose[c(5,6)]), col = "pink", lty = 2,lwd=2)
lines( x= c(1,2), y = exp(Estimatesclose[c(7,8)]), col = "blue", lty = 1,lwd =2)


## plot out of nest from Estimates as reaction norm
plot(rep(c(1,2),4), Estimatesout,  pch = c(21,21,23,23,21,21,23,23), cex = c(2, 2, 2, 2, 1, 1, 1, 1), ylim=c(0, round(max(Estimatesout) + 0.1*max(Estimatesout),0)), xlim = c(0.75,2.25), xlab = "", xaxt = "n", cex.lab=1.5, cex.axis=1.25, ylab = "Out of Nest (s/10min)", bg= col.D, col = "black")
axis(1, at = c(1,2), labels = c("Female","Male"), cex.axis=1.5 )
axis(1, at = 1.5, line = 2, labels = c("Intruder Sex"), cex.axis=2,tick = FALSE)
lines( x= c(1,2), y = Estimatesout[c(1,2)], col = "red", lty = 1, lwd=2)
lines( x= c(1,2), y = Estimatesout[c(3,4)], col = "cyan", lty = 2, lwd =2)
lines( x= c(1,2), y = Estimatesout[c(5,6)], col = "pink", lty = 2, lwd =2)
lines( x= c(1,2), y = Estimatesout[c(7,8)], col = "blue", lty = 1, lwd =2)

plot(rep(c(1,2),4), exp(Estimatesroll),  pch = c(21,21,23,23,21,21,23,23), cex = c(2, 2, 2, 2, 1, 1, 1, 1), ylim=c(0, round(max(exp(Estimatesroll)) + 0.1*max(exp(Estimatesroll)),0)), cex.axis=1.25, cex.lab=1.25, xlim = c(0.75,2.25), xlab = "", xaxt = "n", ylab = "Lateral Rolls (#/10min)", bg= col.D, col = "black")
axis(1, at = c(1,2), labels = c("Female","Male"),cex.axis=1.5 )
axis(1, at = 1.5, line = 2, labels = c("Intruder Sex"), cex.axis=2,tick = FALSE)
# ? lines # there must be a better way but this hack will work
lines( x= c(1,2), y = exp(Estimatesroll[c(1,2)]), col = "red", lty = 1,lwd=2)
lines( x= c(1,2), y = exp(Estimatesroll[c(3,4)]), col = "cyan", lty = 2,lwd=2)
lines( x= c(1,2), y = exp(Estimatesroll[c(5,6)]), col = "pink", lty = 2,lwd=2)
lines( x= c(1,2), y = exp(Estimatesroll[c(7,8)]), col = "blue", lty = 1,lwd =2)
points(x=c(2.2,2.2,2.2,2.2),y=c(2.9,2.6,2.3,2.0),cex=c(2,2,1,1),pch=c(21,23,21,23),bg=c("red","cyan","pink","blue"))
dev.off()

```


````{r}
mod1<-glmer(Int.attack~pairtype*int.sex+(1|order)+(1|IntruderID)+(1|pairID),family = "poisson",df[which(df$size=="large"),])
summary(mod1)
````

Looks like an interaction, like the graphs suggest

Let's make an order effect for the intruder as well as size

Hmmm... can't find dates to put intruder in order... Intruder Size will have to do

````{r}
int.size<-c()
for(i in 1:length(df$Int.SL))
  if(df$Int.SL[i]>=40)
  {
  cat<-"large"
  int.size<-c(int.size,cat)
  }else
  {
  cat<-"small"
  int.size<-c(int.size,cat)
  }
df$int.size<-int.size
mod2<-glmer(Int.attack~pairtype*int.sex*int.size+(1|order)+(1|IntruderID)+(1|pairID),family = "poisson",df[which(df$size=="large"),])
summary(mod2)
````

Still that interaction but nothing else

Looks like this mod is doing worse... no effect of intruder size

````{r}
anova(mod1,mod2)
````

Yep, no difference in model so less complicated should be chosen...

```{r}
newdf<-df[which(df$size=="large"),]
newdf$int.label<-as.factor(paste0(newdf$pairtype,".",newdf$int.sex))
mod6<-glmer(Int.attack~pairtype*int.sex+aggression.C+(1|IntruderID)+(1|pairID),family = "poisson",control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)), df[which(df$size=="large"),])
summary(mod6)
````

````{r}
newdf<-df[which(df$size=="large"),]
newdf$int.label<-as.factor(paste0(newdf$pairtype,".",newdf$int.sex))
mod7<-glmer(Int.attack~int.label+aggression.C+(1|IntruderID)+(1|pairID),family="poisson",newdf)
newdf$int.label<-factor(newdf$int.label, levels=c("femaleLarger.male","maleLarger.female","maleLarger.male","femaleLarger.female"))
mod8<-glmer(Int.attack~int.label+aggression.C+(1|IntruderID)+(1|pairID),family="poisson",newdf)
newdf$int.label<-factor(newdf$int.label, levels=c("maleLarger.female","maleLarger.male","femaleLarger.female","femaleLarger.male"))
mod9<-glmer(Int.attack~int.label+aggression.C+(1|IntruderID)+(1|pairID),family="poisson",newdf)
summary(mod7)
summary(mod8)
summary(mod9)
````

            ML.male   ML.female   FL.male   FL.female

ML.male     --

ML.female   0.0123    --

FL.male     0.0035    0.8780      --

FL.female   0.5285    0.0457      0.0199    --

So... intruders reciprocate aggression from the larger fish. And are more aggressive to opposite sex intruders!

```{r}
Codes <- matrix(0,nrow=4,ncol=4) #8 conditions
Codes[,1] <- rep(1,4)
Codes[2:4,2:4] <- diag(3)
M <- vcov(mod9)[c(1:4),c(1:4)]
FE <- fixef(mod9)[c(1:4)]
Estimates <- as.numeric(Codes %*% FE)
EstVar <- rep(0,length(Estimates))
#
for(i in 1:4) EstVar[i] <- as.numeric(Codes[i,] %*% M %*% t(Codes)[,i])
EstSE <- sqrt(EstVar)
 
 
## plot ckose divider from Estimates as reaction norm
setEPS()
postscript("~/Intruder.eps",print.it = TRUE)
par(mar=c(5,6,1,4))
plot(c(1,1,2,2), exp(Estimates),  pch = c(21,23,21,23), cex = 3, ylim=c(0, round(max(exp(Estimates)) + 0.1*max(exp(Estimates)),0)), cex.axis=1.5,cex.lab=2,xlim = c(0.75,2.25), xlab = "", xaxt = "n", ylab = "Intruder Bites (per/10 min)", bg= c("orange","darkgreen","orange","darkgreen"), col = "black")
axis(1, at = c(1,2), cex.axis=1.5, labels = c("Male Larger","Female Larger") )
axis(1, at = 1.5, line = 2, cex.axis=2, labels = c("Pair Type"), tick = FALSE)
# ? lines # there must be a better way but this hack will work
lines( x= c(1,2), y = exp(Estimates[c(1,3)]), col = "orange", lty = 1,lwd =2)
lines( x= c(1,2), y = exp(Estimates[c(2,4)]), col = "darkgreen", lty = 1,lwd=2)
points(x=c(2.25,2.25),y=c(13.75,12.25),cex=2,pch=c(21,23),bg=c("orange","darkgreen"))
dev.off()
```