---
title: "Intruder Attack"
author: "Drew"
date: "11/02/2022"
output:
  word_document: default
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown of Analysis started 20230515
Start off with all of the items needed in previous analyses
and port in the csv

```{r libraries}
library(lme4)
library(lmerTest)
library(ggplot2)

#Data organizing
#####
#Prior to csv creation, I summed hitdivider.C and bite.C in excel to create aggression.C
df.clean<-read.csv("/Users/andersond/Downloads/20230515_relevant_data.csv")
df<-df.clean

# round the states to remove decimal
df$closedivider.S <-  round(df$closedivider.S, digits = 0)
df$outnest.S <- round(df$outnest.S, digits = 0)

df$order <- as.factor(df$order)
# code for stats and plots for Dennis et al 2022 manuscript
# use this line for published script and supply this data file.
# df <- read.csv(file = "Julies_2022_SuppleTable_BehaviorData.csv")  
df$plot.label <- as.factor(paste(df$pairtype,df$int.sex,df$sex, sep="."))
df$plot.label.order.D <- df$plot.label # makes a duplicate column
df$plot.label.order.D <- factor(df$plot.label, levels = levels(df$plot.label)[c(1,3,6,8,5,7,2,4)]) # reorders data and levels for labels

spacing1 <- c(0,1,3,4,7,8,10,11) # used to place boxes in boxplot and data points
spacing2 <- c(0.5,3.5,7.5,10.5) # used for labeling x-axis

col.D = c("red","red","cyan","cyan","pink","pink", "blue","blue") # colors for plot

par(mar = c(10, 5.1, 4.1, 2.1)) 
## The default value for mar is c(5.1, 4.1, 4.1, 2.1) to reset it
```

Let's check the distribution

```{r}
hist(df$Int.attack)
```
Yep, that's a poisson (more or less)

Running the model as it is written in the for the other works

```{r}
##### Do stats with three-way model, order as fixed and ID and intruder ID as random ####
#### Intruder attack #####
mod.intatt<-glmer(Int.attack~size*sex*int.sex+order+(1|ID)+(1|IntruderID), control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)), family="poisson", data = df) ##
summary (mod.intatt)
```

Interactions come out significant but no individual stats

Let's do the post hoc

```{r}
############################
##### Int.attack graphs linear estimate ##### 
levels(df$plot.label)

a.f.F.L <- df$plot.label == "femaleLarger.female.Female"
b.m.F.L <- df$plot.label == "femaleLarger.male.Female"
c.f.M.L <- df$plot.label == "maleLarger.female.Male"
d.m.M.L <- df$plot.label == "maleLarger.male.Male"
e.f.F.S <- df$plot.label == "maleLarger.female.Female"
f.m.F.S <- df$plot.label == "maleLarger.male.Female"
g.f.M.S <- df$plot.label == "femaleLarger.female.Male"
h.m.M.S <- df$plot.label == "femaleLarger.male.Male"

## create treatment codes
Codes <- matrix(0,nrow=8,ncol=8) #8 conditions
Codes[,1] <- rep(1,8)
Codes[2:8,2:8] <- diag(7)
# Codes
col.D = c("red","red","cyan","cyan","pink","pink", "blue","blue")
BarSpace <- c(0.5,1,2,1,3,1,2,1) #redo the codes
#####
glm2plot <- glmer(data = df, df$Int.attack~ b.m.F.L + c.f.M.L + d.m.M.L + e.f.F.S + f.m.F.S + g.f.M.S + h.m.M.S + order+ (1|ID)+(1|IntruderID), family = "poisson") ## 

summary(glm2plot) 
# the following does the pairwise comparisons that aren't in the initial model
#two cyan (c vs d)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,-1,0,0,0,0,0,0,0))         
# two pink (e vs f) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,1,-1,0,0,0,0,0))
# two blue (g vs h)  
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,0,0,1,-1,0,0,0))
# two red (a vs b) 
#this value is from summary(glm2plot)

# Sex (larger w female.int)  (a vs c) see above
#this value is from summary(glm2plot)
# Sex (larger w male.int)  (b vs d) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,-1,0,0,0,0,0,0,0))
# Sex (smaller w female.int) (e vs g) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,1,0,-1,0,0,0,0))
# Sex (smaller w male.int) (f vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,0,0,1,0,-1,0,0,0))

# Size (Female w male.int) (b vs f) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,0,0,-1,0,0,0,0,0))
# Size (Male w female.int) (c vs g) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,0,0,0,-1,0,0,0,0))
# Size (Male w male.int) (d vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,1,0,0,0,-1,0,0,0))
# Size (Female w female.int) (a vs e) see above

##new
# Size and sex (femalelarger w female.int) (a vs g) is above
# Size and sex (Femalelarger w male.int) (b vs h) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,1,0,0,0,0,0,-1,0,0,0))
# Size and sex (Malelarger w female.int) (c vs e) 
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,1,0,-1,0,0,0,0,0,0)) 
# Size and sex (Malelarger w male.int) (d vs f) (0.2875)
linearHypothesis(glm2plot, coef=fixef(glm2plot), hypothesis.matrix = c(0,0,0,1,0,-1,0,0,0,0,0))
```
a:b NS       two red??

c:d 0.00511  two cyan??

e:f 0.00511  two pink??

g:h NS       two blue??

b:d 0.0141   larger w female.int??

e:g NS.      smaller w female.int??

f:h 0.0141   smaller w male.int??

b:f 0.0141   Female w male.int??

c:g NS.      Male w female.int??

d:h 0.0141.  Male w male.int??

b:h NS       Femalelarger w male.int??

c:e NS       Malelarger w female.int??

d:f NS       Malelarger w male.int??

###
!!!!
Ry, need to explain these comparisons better!
!!!!

So that's weird that our p-values are the same across so many of these

Let's plot

```{r}
M <- vcov(glm2plot)[c(1:8),c(1:8)]
FE <- fixef(glm2plot)[c(1:8)]
Estimates <- as.numeric(Codes %*% FE)
EstVar <- rep(0,length(Estimates))
#
for(i in 1:8) EstVar[i] <- as.numeric(Codes[i,] %*% M %*% t(Codes)[,i])
EstSE <- sqrt(EstVar)


## plot ckose divider from Estimates as reaction norm
plot(rep(c(1,2),4), exp(Estimates),  pch = c(21,21,23,23,21,21,23,23), cex = c(2, 2, 2, 2, 1, 1, 1, 1), ylim=c(0, round(max(exp(Estimates)) + 0.1*max(exp(Estimates)),0)), xlim = c(0.75,2.25), xlab = "", xaxt = "n", ylab = "Intruder Attacks (per/10 min)", bg= col.D, col = "black")
axis(1, at = c(1,2), labels = c("female","male") )
axis(1, at = 1.5, line = 2, labels = c("Intruder Sex"), tick = FALSE)
# ? lines # there must be a better way but this hack will work
lines( x= c(1,2), y = exp(Estimates[c(1,2)]), col = "red", lty = 1)
lines( x= c(1,2), y = exp(Estimates[c(3,4)]), col = "cyan", lty = 2)
lines( x= c(1,2), y = exp(Estimates[c(5,6)]), col = "pink", lty = 2)
lines( x= c(1,2), y = exp(Estimates[c(7,8)]), col = "blue", lty = 1)

```

Well...

Ah!

Of course!

The intruder attack is not based on individual fish in the pair, it's based on the pair itself. So including the size of sex or sex doesn't matter. What we should be checking is the pair type.

Let's just do a simple check

How does pairtype and intruder sex affect intruder attacks?

To avoid double sampling just sample one size of the pair

````{r}
boxplot(Int.attack~pairtype*int.sex,df[which(df$size=="large"),])
````

Hmmmm... looks similar to what we saw previously

What about straight up correlation of the attacks to attacks?

````{r}
plot(Int.attack~aggression.C,df)
````

Not looking good, but let's just consider the larger fish in the pair

````{r}
plot(Int.attack~aggression.C,df[which(df$size=="large"),])
````

That's better!

Let's do a simple model

````{r}
mod<-lm(Int.attack~aggression.C,df[which(df$size=="large"),])
summary(mod)
````

Alright. Let's do some better modeling

We don't need order, because that's the order effect of the pair, not the intruder

````{r}
mod1<-glmer(Int.attack~pairtype*int.sex+(1|order)+(1|IntruderID)+(1|pairID),family = "poisson",df[which(df$size=="large"),])
summary(mod1)
````

Looks like an interaction, like the graphs suggest

Let's make an order effect for the intruder as well as size

Hmmm... can't find dates to put intruder in order... Intruder Size will have to do

````{r}
int.size<-c()
for(i in 1:length(df$Int.SL))
  if(df$Int.SL[i]>=40)
  {
  cat<-"large"
  int.size<-c(int.size,cat)
  }else
  {
  cat<-"small"
  int.size<-c(int.size,cat)
  }
df$int.size<-int.size
mod2<-glmer(Int.attack~pairtype*int.sex*int.size+(1|order)+(1|IntruderID)+(1|pairID),family = "poisson",df[which(df$size=="large"),])
summary(mod2)
````

Still that interaction but nothing else

Looks like this mod is doing worse... no effect of intruder size

````{r}
anova(mod1,mod2)
````

Yep, no difference in model so less complicated should be chosen...

```{r}
newdf<-df[which(df$size=="large"),]
newdf$int.label<-as.factor(paste0(newdf$pairtype,".",newdf$int.sex))
mod3<-glmer(Int.attack~int.label+(1|order)+(1|IntruderID)+(1|pairID),family="poisson",newdf)
newdf$int.label<-factor(newdf$int.label, levels=c("femaleLarger.male","maleLarger.female","maleLarger.male","femaleLarger.female"))
mod4<-glmer(Int.attack~int.label+(1|order)+(1|IntruderID)+(1|pairID),family="poisson",newdf)
newdf$int.label<-factor(newdf$int.label, levels=c("maleLarger.female","maleLarger.male","femaleLarger.female","femaleLarger.male"))
mod5<-glmer(Int.attack~int.label+(1|order)+(1|IntruderID)+(1|pairID),family="poisson",newdf)
summary(mod3)
summary(mod4)
summary(mod5)
````

We can see in mod5 that the big driver is the male to female comparison in the male larger pair with Male vs. male aggression driving it. I believe this supports our claim of Male vs Male aggression

Let's do a linear variable

````{r}
mod6<-glmer(Int.attack~pairtype*int.sex*aggression.C+(1|order)+(1|IntruderID)+(1|pairID),family = "poisson",control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)), df[which(df$size=="large"),])
summary(mod6)
````

This is too much to model (see the Singularity warning!!) let's drop order

````{r}
mod6<-glmer(Int.attack~pairtype*int.sex*aggression.C+(1|IntruderID)+(1|pairID),family = "poisson",control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)), df[which(df$size=="large"),])
summary(mod6)
````

STILL too much

Let's just drop the interaction of aggression

````{r}
mod6<-glmer(Int.attack~pairtype*int.sex+aggression.C+(1|order)+(1|IntruderID)+(1|pairID),family = "poisson",control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)), df[which(df$size=="large"),])
summary(mod6)
````

Whomp whomp

Drop order

````{r}
mod6<-glmer(Int.attack~pairtype*int.sex+aggression.C+(1|IntruderID)+(1|pairID),family = "poisson",control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)), df[which(df$size=="large"),])
summary(mod6)
````

Well looky here! We have everything significant with a big amount from aggression and interaction!!

Let's try this!

````{r}
newdf<-df[which(df$size=="large"),]
newdf$int.label<-as.factor(paste0(newdf$pairtype,".",newdf$int.sex))
mod7<-glmer(Int.attack~int.label+aggression.C+(1|IntruderID)+(1|pairID),family="poisson",newdf)
newdf$int.label<-factor(newdf$int.label, levels=c("femaleLarger.male","maleLarger.female","maleLarger.male","femaleLarger.female"))
mod8<-glmer(Int.attack~int.label+aggression.C+(1|IntruderID)+(1|pairID),family="poisson",newdf)
newdf$int.label<-factor(newdf$int.label, levels=c("maleLarger.female","maleLarger.male","femaleLarger.female","femaleLarger.male"))
mod9<-glmer(Int.attack~int.label+aggression.C+(1|IntruderID)+(1|pairID),family="poisson",newdf)
summary(mod7)
summary(mod8)
summary(mod9)
````

Mwahahahah!!

            ML.male   ML.female   FL.male   FL.female

ML.male     --

ML.female   0.0123    --

FL.male     0.0035    0.8780      --

FL.female   0.5285    0.0457      0.0199    --

So... intruders reciprocate aggression from the larger fish. And are more aggressive to opposite sex intruders!

Weirdly it seems there is a stronger response in FvF than MvM (according to mod6)

````{r}
plot(Int.attack~int.label+aggression.C,newdf)
````

There it is! That is there... but it's mostly due to those high instances in the female intruder


```{r}
Codes <- matrix(0,nrow=4,ncol=4) #8 conditions
Codes[,1] <- rep(1,4)
Codes[2:4,2:4] <- diag(3)
M <- vcov(mod9)[c(1:4),c(1:4)]
FE <- fixef(mod9)[c(1:4)]
Estimates <- as.numeric(Codes %*% FE)
EstVar <- rep(0,length(Estimates))
#
for(i in 1:4) EstVar[i] <- as.numeric(Codes[i,] %*% M %*% t(Codes)[,i])
EstSE <- sqrt(EstVar)
 
 
## plot ckose divider from Estimates as reaction norm
setEPS()
postscript("/Users/andersond/Desktop/Intruder.eps",print.it = TRUE)
par(mar=c(5,6,1,4))
plot(c(1,1,2,2), exp(Estimates),  pch = c(21,23,21,23), cex = 3, ylim=c(0, round(max(exp(Estimates)) + 0.1*max(exp(Estimates)),0)), cex.axis=1.5,cex.lab=2,xlim = c(0.75,2.25), xlab = "", xaxt = "n", ylab = "Intruder Bites (per/10 min)", bg= c("orange","darkgreen","orange","darkgreen"), col = "black")
axis(1, at = c(1,2), cex.axis=1.5, labels = c("Male Larger","Female Larger") )
axis(1, at = 1.5, line = 2, cex.axis=2, labels = c("Pair Type"), tick = FALSE)
# ? lines # there must be a better way but this hack will work
lines( x= c(1,2), y = exp(Estimates[c(1,3)]), col = "orange", lty = 1,lwd =2)
lines( x= c(1,2), y = exp(Estimates[c(2,4)]), col = "darkgreen", lty = 1,lwd=2)
points(x=c(2.25,2.25),y=c(13.75,12.25),cex=2,pch=c(21,23),bg=c("orange","darkgreen"))
dev.off()
```